Development and Implementation of an ANN Based Flow Law for Numerical Simulations of Thermo-Mechanical Processes at High Temperatures in FEM Software

### 2023-10-16 Algorithms-2023.tex
Rename final production article

%=================================================================
\documentclass[algorithms,article,accept,pdftex,moreauthors]{Definitions/mdpi}
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

@@

%=================================================================
% MDPI internal commands
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2023}
\copyrightyear{2023}
\externaleditor{{ Xiang Zhang, Xiaoxiao Li } %MDPI: please add the academic editors.
}
\datereceived{16 December 2022}
\daterevised{10 January 2023} % Only for the journal Acoustics
\dateaccepted{12 January 2023}
\datepublished{}
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

@@

% Authors, for the paper (add full first names)
\Author{Olivier Pantal\'{e}\orcidA{}}

@@

% Affiliations / Addresses (Add [XX] after \address if there is only one affiliation.)
\address[1]{Laboratoire Génie de Production, Institut National Polytechnique/Ecole Nationale d'Ingénieurs de Tarbes, Université de Toulouse, 47 Av d'Azereix, F-65016 Tarbes, France; Olivier.Pantale@enit.fr; Tel.: +33-562442933}

@@

% Keywords
\keyword{ANN flow law; constitutive behavior; radial return algorithm; numerical implementation; VUHARD ; GrC15; Abaqus Explicit}

@@

In the thermomechanical simulation of forming processes, these behavior laws define the dependence~\cite{Lee-2006} of the flow stress of the material $\sigma^y$ as a function of the three input variables, which are the plastic strain $\varepsilon^p$, the~strain rate $\mdot\varepsilon$ and the temperature $T$ of the material, so that the general form of the flow law can be written with the following expression:
\begin{equation}
\sigma^y = f\left(\varepsilon^p,\mdot\varepsilon,T\right)\label{eq:GenFlow}
\end{equation}

@@

The underlying idea proposed in our approach is to implement a flow law described by a trained ANN as a FORTRAN 77 subroutine in the Abaqus FE code.
This ANN was previously trained from the data extracted from mechanical tests of the material and can directly define the value of the flow stress $\sigma^y$ as a function of the plastic strain $\varepsilon^p$, the~strain rate $\mdot\varepsilon$ and the temperature $T$.
After a training phase based on the use of the Python library TensorFlow~\cite{Abadi-2016, Mattmann-2020}, the~weights and biases of the trained neural network are transcoded into a subroutine in FORTRAN 77, which is compiled and linked with the libraries of the Abaqus FE code to include the behavior of the material by allowing the computation of the flow stress $\sigma^y$ as a function of $\varepsilon^p$, $\mdot\varepsilon$ and $T$, and~of its three derivatives $\partial\sigma^y/\partial\varepsilon^p$, $\partial\sigma^y/\partial\mdot\varepsilon$ and $\partial\sigma^y/\partial T$.

@@

Conforming to the structure of the ANN reported in Figure~\ref{fig:ANN-scheme} any hidden layer $k$, containing $n$ neurons, takes a weighted sum of the outputs $\overrightarrow{\hat{y}}^{(k-1)}$ of the immediately previous layer $(k-1)$, containing $m$ neurons, given by the following equation:
\begin{equation}
y_i\lay{k} = \sum_{j=1}^m w_{ij}\lay{k} \hat{y}_j^{(k-1)}+ b_i\lay{k},\label{eq:ANN1}
\end{equation}
where $y_i\lay{k}$ is the entry of the $i$th neuron of layer $k$, $\hat{y}_j\lay{k-1}$ is the output of the $j$th neuron of layer $(k-1)$, $w_{ij}\lay{k}$ is the associated weight parameter between the $i$th neuron of layer $k$ and the $j$th neuron of layer $(k-1)$ and $b_i\lay{k}$ is the associated bias of the $i$th neuron of layer $k$.
Those weights $w_{ij}$ and bias $b_i$, for~each layer, are the training parameters of the ANN that we have to adjust during the training process.
For the proposed model, we selected the sigmoid activation function so that each neuron in the hidden layer $k$ provides an output value ${\hat{y}}$ from the input value $y$ of the same neuron defined by Equation~(\ref{eq:ANN1}) according to the following equation:
\begin{equation}
\hat{y}=\frac{1}{1 + \e{-y}}\label{eq:ANN2}
\end{equation}

@@

Then we compute the output $s$ of the ANN from the output vector of the second hidden layer $\overrightarrow{y}_2$ using the following equation:
\begin{equation}
s = \overrightarrow{w}^T \dotp \overrightarrow{y}_2 + b\label{eq:ANN5}
\end{equation}

@@

In neural network learning, it is necessary to define the objective function to be minimized and the evaluation of the model error.
In this study, the~error evaluation is based on the mean square error ($\MSE$) and the root mean square error ($\RMSE$) given by the following equation:
\begin{equation}
\RMSE (\MPa) = \sqrt{\MSE} = \sqrt{\frac{1}{N} \sum_{i=1}^{N} \left(\square_i^e - \square_i^y\right)^2}, \label{eq:RMSE}
\end{equation}
where $N$ is the total number of numerical training data used, $\square_i^y$ is the $i$th value predicted by the neural network, and~$\square_i^e$ is the corresponding experimental value coming from the experimental tests.
The accuracy and predictive ability of the models is assessed by the mean absolute relative error ($\MARE$) defined by Equation~(\ref{eq:AARE}):
\begin{equation}
\MARE(\%) = \frac{1}{N} \sum_{i=1}^{N}{\left|\frac{\square_i^y -\square_i^e}{\square_i^e}\right|} \times 100 \label{eq:AARE}
\end{equation}

@@

The set of these data is used as training data for the neural network.
Several neural network architectures have been studied in this work; they differ from each other by the number of neurons present in the two hidden layers.
Among them, we selected 4 different architectures named 3-7-4-1, 3-9-4-1, 3-9-7-1 and 3-15-7-1 for which the name 3-$n$-$m$-1 translates an ANN with 2 hidden layers, having $n$ neurons on the first layer and $m$ neurons on the second~layer.

All models have been trained for the same number of iterations (50,000 iterations), and~around $50$ min of training on a Dell XPS-13 7390 laptop running Ubuntu 22.04 LTS 64 bits with 16 GB of RAM and an Intel 4-core i7-10510U processor allow obtaining the converged parameters of the ANN models.
Figure~\ref{fig:Convergence-ANN} shows the evolution of the training error defined by the $\log_{10}$ of the mean square error ($\log_{10}\left[\MSE\right]$) during the training~phase.
\begin{figure}[H]
\includegraphics[width=0.65\columnwidth]{Figures/Convergence-ANN}
\caption{Convergence of the ANN models during the training~phase.}
\label{fig:Convergence-ANN}
\end{figure}
As we can see on this figure, after~50,000 iterations, we can consider that we have reached a stationary state of the model learning and that it is useless to continue the learning phase.
As expected, the~more neurons the model contains, the~more it can follow the non-linear evolution of the material’s behavior and therefore the more the mean square error ($\MSE$) during the learning phase decreases.
Table~\ref{tab:ErrorANN} shows the main results of the training of these neural~networks.
\begin{table}[H]
\caption{Results concerning the training of the four ANN flow~laws. \label{tab:ErrorANN}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{lCCCCC}
	\toprule
	\textbf{ANN}      & \boldmath{$n_v$} &  \boldmath{$t$}  &     {\bf $\MSE$}      & {\bf $\MARE$} & {\bf $\RMSE$} \\
	         &       & \textbf{(min)} & \boldmath{$\times 10^{-5}$} &  \textbf{(\%)}   &  \textbf{(MPa)}  \\ \midrule
	3-7-4-1  &  65   &  48   &       3.91       &  1.88   &  3.05   \\
	3-9-4-1  &  81   &  48   &       3.29       &  1.70   &  2.75   \\
	3-9-7-1  &  114  &  49   &       1.83       &  1.25   &  2.44   \\
	3-15-7-1 &  180  &  50   &       1.01       &  0.97   &  2.30   \\ \bottomrule
\end{tabularx}

@@

Concerning the performance of the ANN flow laws, the~correlation results for both reported models, are much better than the ones obtained by Zhou et~al.~\cite{Zhou-2020} during his work on the same material with four different analytical flow laws, especially since he had to split the data into two groups according to the temperature value (one on the range \mbox{$T=750-850~\celsius$} and one on the range \mbox{$T=850-1300~\celsius$}) and to identify two sets of parameters for each flow law to reduce the error of his identified analytical models.
\textls[-11]{This of course raises the question of the usability of those analytical laws where the temperature of the material changes from one group to the other during a thermomechanical~transformation.}

In our approach and by using an ANN flow law, the~identified law is not only valid over the whole temperature range, but~it displays a $\MARE$ value $5$ times lower than the best flow law proposed by Zhou et~al.~\cite{Zhou-2020}: the Arrhenius law with an $\MARE=3.74\%$ over the range $T=750-850~\celsius$ and $\MARE=5.76\%$ over the range $T=850-1300~\celsius$, while the $\MARE=0.97\%$ for the 3-15-7-1 and $\MARE=1.88\%$ for the 3-7-4-1 ANN flow laws proposed~here.

The disadvantage of developing a flow law model based on neural networks is the number of internal variables in the network ($180$ in the case of the 3-15-7-1 network), which makes it difficult to translate the network into printable results.
Using a Johnson--Cook-type flow law, for~example, allows the reader to quickly get an idea of the law, where the analytical formulation of the law is known to the users, and~the behavior of a material is based on the knowledge of only $5$ internal parameters to be identified, which makes it easy to publish in a table.
Concerning an Arrhenius law, this task becomes a little more complex, as one can have from $24$ to $36$ coefficients.
However, in our case, the~publication of the $65$ coefficients of the 3-7-4-1 model or the $180$ coefficients of the 3-15-7-1 model makes this task delicate.
As an illustration, we provide in the Appendix \ref{sec:Appendix} all the coefficients of the 3-7-4-1 model identified during this~study.

@@

Once the neural network is trained as presented in Section~\ref{sec:ANN-traning}, it can be used in a finite element code for the numerical simulation of a structure subjected to thermomechanical loading.
This requires the extraction of the internal variables of the neural network and their transfer as a subroutine in FORTRAN 77 based on equations proposed in Sections~\ref{sec:ANN-equations} and \ref{sec:ANN-derivative}.

@@

If we refer to the general flowchart of a finite element code as shown in Figure~\ref{fig:StressUpdate}, integrating the flow law described by the ANN concerns the computation of the stress tensor $\Sig_1$ at the end of an increment, in~the yellow rectangle on the~figure.
\begin{figure}[H]
\includegraphics[width=0.9\columnwidth]{Figures/StressUpdate}
\caption{General flowchart of a FE code, focus on the stresses~computation using an iterative solving procedure.}
\label{fig:StressUpdate}
\end{figure}
Within the framework of a FE formulation in large deformations such as the one used in thermomechanical modeling of processes, this computation of the stress tensor $\Sig_1$ is to be carried out on all the integration points of each element of the studied structure.
Since the numerical model can include thousands of elements, themselves comprising between $1$ and $8$ integration points depending on the elements used, this stress computation must be as fast as possible in order not to increase the CPU time too much, but~precise enough so that the results are under the physics of the process.
This is even more important if we want to integrate this flow law in an explicit FE code, such as Abaqus/Explicit, for which one second of physical simulation corresponds to several million iterations of these stress computations.
Thus, the~complexity of the ANN, \ie the number of computational steps that must be performed to compute the flow stress as a function of the input variables, is a major parameter in the choice of the neural network.
As an example, for~the model presented in Section~\ref{sec:ANN-traning}, $180$ internal variables, $15$ neurons on the first hidden layer and $7$ neurons on the second hidden layer were listed.
Given the equations described in Sections~\ref{sec:ANN-equations} and~\ref{sec:ANN-derivative}, it will be necessary to compute $22$ exponentials, to~make matrix--vector products of size $15\times3$ and $15\times7$ plus many other numerical operations to compute the flow stress $\sigma^y$ and the $3$ derivatives of it with respect to $\varepsilon^p$, $\mdot\varepsilon$ and $T$.

Implementing the ANN flow law identified above is realized here in a VUHARD subroutine, similarly as proposed by van Rensburg et~al.~\cite{JansenVanRensburg-2012}, used by the Abaqus Explicit FE code in order to allow a user to program the computation of the flow stress $\sigma^y$ and its 3 derivatives as a function of the model input data.
This subroutine is used when calculating the stress tensor $\Sig_1$ at the end of an increment from the stress tensor at the beginning of the increment $\Sig_0$, the~deformations, the~material parameters and the history of the deformation at each finite element integration points, according to the stress integration algorithm based on the radial return method as described in Simo et~al.~\cite{Simo-1998} for the general aspects, Ming et~al.~\cite{Ming-2018} for Abaqus Explicit FE code, or Pantalé et~al.~\cite{Pantale-2004} for the DynELA FE code.
Thus, without~going into too much detail about the stress integration scheme used in finite element codes (the curious reader can refer to~\cite{Ponthot-2002, Ming-2018, Pantale-2004, Liang-2022} for details about this method), Figure~\ref{fig:RadialReturn} shows the location of the VUHARD subroutine used to compute the flow stress $\sigma^y$ and its derivative $\partial\sigma^y/\partial\Gamma$ used in the writing of the two quantities $\gamma(\Gamma)$ and $\gamma^{'}(\Gamma)$ used in the Newton--Raphson solving procedure from the following relation:
\begin{equation}
\frac{d\sigma^{y}}{d\Gamma} = \sqrt{\frac{2}{3}}\left(\frac{\partial\sigma^{y}}{\partial\varepsilon^{p}}+\frac{1}{\Delta t}\frac{\partial\sigma^{y}}{\partial\mdot{\varepsilon}^p}+\frac{\eta\sigma^{y}}{\rho C_{p}}\frac{\partial\sigma^{y}}{\partial T}\right)\label{eq:derivatives},
\end{equation}
where $\Gamma$ is the consistency parameter used in the radial return algorithm as defined by Simo et~al.~\cite{Simo-1998}, $\Delta t$ is the time increment, $\eta$ is the Taylor--Quinney coefficient defining the amount of plastic work converted into heat energy, $C_{p}$ is the specific heat coefficient, $\rho$ is the density of the material and $\partial \sigma^y/\partial\varepsilon^p$, $\partial \sigma^y/\partial\mdot\varepsilon$ and $\partial \sigma^y/\partial T$ are the three derivatives defined in Equation~(\ref{eq:DANN3}).
\begin{figure}[H]
\includegraphics[width=0.6\columnwidth]{Figures/VumatGeneralNewtonRaphson}
\caption{General flowchart of the radial return algorithm to compute the final stress tensor $\Sig_1$.}
\label{fig:RadialReturn}
\end{figure}
\unskip

@@

The mesh of the sample is made with $850$ axis-symmetric quadrilateral finite elements with $4$ nodes and reduced integration (named CAX4R in the Abaqus software) with $50$ elements in the vertical direction and $17$ elements in the radial direction, respectively.
The cylinder is between two rigid surfaces, and the Coulomb friction law with a friction coefficient at the contact surfaces was set to $\mu=0.15$.
The simulation time being fixed at $10~\text{s}$ in~order to reduce the simulation time, considering that an explicit integration scheme is used, a~global mass scaling is used for all simulations.
The VUHARD subroutine is compiled using the GNU gfortran 11.3.0 and linked to the main Abaqus Explicit executable.
All benchmarks tests were solved using Abaqus Explicit 2022 on a Dell XPS 13 laptop running Ubuntu 20.04 64~bits with 16~GiB of RAM and one 4 core i7-10510U Intel Processor.
All computations were performed using the double precision option of Abaqus, with~parallel threads execution on two~cores.

>>>>

%=================================================================
\documentclass[algorithms,article,accept,pdftex,oneauthors]{Definitions/mdpi}
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

@@

%=================================================================
% MDPI internal commands
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{16}
\issuenum{1}
\articlenumber{56}
\pubyear{2023}
\copyrightyear{2023}
\externaleditor{Academic Editors:{ Xiang Zhang and Xiaoxiao Li } %MDPI: please add the academic editors.
}
\datereceived{16 December 2022}
\daterevised{10 January 2023} % Only for the journal Acoustics
\dateaccepted{12 January 2023}
\datepublished{13 January 2023}
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/10.3390/\linebreak a16010056} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

@@

% Authors, for the paper (add full first names)
\Author{{Olivier Pantal\'{e}} %MDPI: Please carefully check the accuracy of names and affiliations.
% Name is correct
 \orcidA{}}

@@

% Affiliations / Addresses (Add [XX] after \address if there is only one affiliation.)
\address[1]{Laboratoire Génie de Production, Institut National Polytechnique/Ecole Nationale d'Ingénieurs de Tarbes, Université de Toulouse, 47 Av d'Azereix, F-65016 Tarbes, France; olivier.pantale@enit.fr; Tel.: +33-562442933}

@@

% Keywords
\keyword{ANN flow law; constitutive behavior; radial return algorithm; numerical implementation; VUHARD; GrC15; Abaqus Explicit}

@@

In the thermomechanical simulation of forming processes, these behavior laws define the dependence~\cite{Lee-2006} of the flow stress of the material $\sigma^y$ as a function of the three input variables, which are the plastic strain $\varepsilon^p$, the~strain rate $\mdot\varepsilon$ and the temperature $T$ of the material, so that the general form of the flow law can be written with the following~expression:
\begin{equation}
\sigma^y = f\left(\varepsilon^p,\mdot\varepsilon,T\right)\label{eq:GenFlow}
\end{equation}

@@

The underlying idea proposed in our approach is to implement a flow law described by a trained ANN as a FORTRAN 77 subroutine in the Abaqus FE code.
This ANN was previously trained from the data extracted from mechanical tests of the material and can directly define the value of the flow stress $\sigma^y$ as a function of the plastic strain $\varepsilon^p$, the~strain rate $\mdot\varepsilon$ and the temperature $T$.
After a training phase based on the use of the Python library TensorFlow~\cite{Abadi-2016, Mattmann-2020}, the~weights and biases of the trained neural network are transcoded into a subroutine in FORTRAN 77, which is compiled and linked with the libraries of the Abaqus FE code to include the behavior of the material by allowing the computation of the flow stress $\sigma^y$ as a function of $\varepsilon^p$, $\mdot\varepsilon$ and $T$, and~of its three derivatives $\partial\sigma^y/\partial\varepsilon^p$, $\partial\sigma^y/\partial\mdot\varepsilon$ and~$\partial\sigma^y/\partial T$.

@@

Conforming to the structure of the ANN reported in Figure~\ref{fig:ANN-scheme} any hidden layer $k$, containing $n$ neurons, takes a weighted sum of the outputs $\overrightarrow{\hat{y}}^{(k-1)}$ of the immediately previous layer $(k-1)$, containing $m$ neurons, given by the following equation:
\begin{equation}
y_i\lay{k} = \sum_{j=1}^m w_{ij}\lay{k} \hat{y}_j^{(k-1)}+ b_i\lay{k},\label{eq:ANN1}
\end{equation}
where $y_i\lay{k}$ is the entry of the $i${{th}} %MDPI: OK for me
 neuron of layer $k$, $\hat{y}_j\lay{k-1}$ is the output of the $j${{th}} neuron of layer $(k-1)$, $w_{ij}\lay{k}$ is the associated weight parameter between the $i${{th}} neuron of layer $k$ and the $j${{th}} neuron of layer $(k-1)$ and $b_i\lay{k}$ is the associated bias of the $i${{th}} neuron of layer $k$.
Those weights $w_{ij}$ and bias $b_i$, for~each layer, are the training parameters of the ANN that we have to adjust during the training process.
For the proposed model, we selected the sigmoid activation function so that each neuron in the hidden layer $k$ provides an output value ${\hat{y}}$ from the input value $y$ of the same neuron defined by Equation~(\ref{eq:ANN1}) according to the following equation:
\begin{equation}
\hat{y}=\frac{1}{1 + \e{-y}}\label{eq:ANN2}
\end{equation}

@@

{Then} %MDPI: This is OK for me
 we compute the output $s$ of the ANN from the output vector of the second hidden layer $\overrightarrow{y}_2$ using the following equation:
\begin{equation}
s = \overrightarrow{w}^T \dotp \overrightarrow{y}_2 + b\label{eq:ANN5}
\end{equation}

@@

In neural network learning, it is necessary to define the objective function to be minimized and the evaluation of the model error.
In this study, the~error evaluation is based on the mean square error ($\MSE$) and the root mean square error ($\RMSE$) given by the following equation:
\begin{equation}
\RMSE (\MPa) = \sqrt{\MSE} = \sqrt{\frac{1}{N} \sum_{i=1}^{N} \left(\square_i^e - \square_i^y\right)^2}, \label{eq:RMSE}
\end{equation}
where $N$ is the total number of numerical training data used, $\square_i^y$ is the $i${{th}} value predicted by the neural network, and~$\square_i^e$ is the corresponding experimental value coming from the experimental tests.
The accuracy and predictive ability of the models is assessed by the mean absolute relative error ($\MARE$) defined by Equation~(\ref{eq:AARE}):
\begin{equation}
\MARE(\%) = \frac{1}{N} \sum_{i=1}^{N}{\left|\frac{\square_i^y -\square_i^e}{\square_i^e}\right|} \times 100 \label{eq:AARE}
\end{equation}

@@

The set of these data is used as training data for the neural network.
Several neural network architectures have been studied in this work; they differ from each other by the number of neurons present in the two hidden layers.
Among them, we selected 4~different architectures named 3-7-4-1, 3-9-4-1, 3-9-7-1 and 3-15-7-1 for which the name 3-$n$-$m$-1 translates an ANN with 2 hidden layers, having $n$ neurons on the first layer and $m$ neurons on the second~layer.

All models have been trained for the same number of iterations (50,000 iterations), and~around $50$ min of training on a Dell XPS-13 7390 laptop running Ubuntu 22.04 LTS 64 bits with 16 GB of RAM and an Intel 4-core i7-10510U processor allow obtaining the converged parameters of the ANN models.
Figure~\ref{fig:Convergence-ANN} shows the evolution of the training error defined by the $\log_{10}$ of the mean square error ($\log_{10}\left[\MSE\right]$) during the training~phase.
\begin{figure}[H]
\includegraphics[width=0.65\columnwidth]{Figures/Convergence-ANN}
\caption{{Convergence} %MDPI: This has been done here
 of the ANN models during the training~phase.}
\label{fig:Convergence-ANN}
\end{figure}
As we can see on this figure, after~50,000 iterations, we can consider that we have reached a stationary state of the model learning and that it is useless to continue the learning phase.
As expected, the~more neurons the model contains, the~more it can follow the non-linear evolution of the material’s behavior and therefore the more the mean square error ($\MSE$) during the learning phase decreases.
Table~\ref{tab:ErrorANN} shows the main results of the training of these neural~networks.
\begin{table}[H]
\caption{Results concerning the training of the four ANN flow~laws. \label{tab:ErrorANN}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{lCCCCC}
	\toprule
	\textbf{ANN}      & \boldmath{$n_v$} &  \boldmath{$t$}  &     {\bf $\MSE$}      & {\bf $\MARE$} & {\bf $\RMSE$} \\
	         &       & \textbf{(min)} & \boldmath{$\times 10^{-5}$} &  \textbf{(\%)}   &  \textbf{(MPa)}  \\ \midrule
	3-7-4-1  &  65   &  48   &       3.91       &  1.88   &  3.05   \\
	3-9-4-1  &  81   &  48   &       3.29       &  1.70   &  2.75   \\
	3-9-7-1  &  114  &  49   &       1.83       &  1.25   &  2.44   \\
	3-15-7-1 &  180  &  50   &       1.01       &  0.97   &  2.30   \\ \bottomrule
\end{tabularx}

@@

Concerning the performance of the ANN flow laws, the~correlation results for both reported models, are much better than the ones obtained by Zhou et~al.~\cite{Zhou-2020} during his work on the same material with four different analytical flow laws, especially since he had to split the data into two groups according to the temperature value (one on the range \mbox{$T=750$--$850~\celsius$} and one on the range \mbox{$T=850$--$1300~\celsius$}) and to identify two sets of parameters for each flow law to reduce the error of his identified analytical models.
\textls[-11]{This of course raises the question of the usability of those analytical laws where the temperature of the material changes from one group to the other during a thermomechanical~transformation.}

In our approach and by using an ANN flow law, the~identified law is not only valid over the whole temperature range, but~it displays a $\MARE$ value $5$ times lower than the best flow law proposed by Zhou et~al.~\cite{Zhou-2020}: the Arrhenius law with an $\MARE=3.74\%$ over the range $T=750$--$850~\celsius$ and $\MARE=5.76\%$ over the range $T=850$--$1300~\celsius$, while the $\MARE=0.97\%$ for the 3-15-7-1 and $\MARE=1.88\%$ for the 3-7-4-1 ANN flow laws proposed~here.

The disadvantage of developing a flow law model based on neural networks is the number of internal variables in the network ($180$ in the case of the 3-15-7-1 network), which makes it difficult to translate the network into printable results.
Using a Johnson--Cook-type flow law, for~example, allows the reader to quickly get an idea of the law, where the analytical formulation of the law is known to the users, and~the behavior of a material is based on the knowledge of only $5$ internal parameters to be identified, which makes it easy to publish in a table.
Concerning an Arrhenius law, this task becomes a little more complex, as one can have from $24$ to $36$ coefficients.
However, in our case, the~publication of the $65$~coefficients of the 3-7-4-1 model or the $180$ coefficients of the 3-15-7-1 model makes this task delicate.
As an illustration, we provide in the Appendix \ref{sec:Appendix} all the coefficients of the 3-7-4-1 model identified during this~study.

@@

\textls[-15]{Once the neural network is trained as presented in Section~\ref{sec:ANN-traning}, it can be used in a finite element code for the numerical simulation of a structure subjected to thermomechanical loading.
This requires the extraction of the internal variables of the neural network and their transfer as a subroutine in FORTRAN 77 based on equations proposed in Sections~\ref{sec:ANN-equations} and \ref{sec:ANN-derivative}.}

@@

If we refer to the general flowchart of a finite element code as shown in Figure~\ref{fig:StressUpdate}, integrating the flow law described by the ANN concerns the computation of the stress tensor $\Sig_1$ {at the end of} %MDPI:\Sig is the stress tensor, while \sigma is a scalar value
 an increment, in~the yellow rectangle on the~figure.
\begin{figure}[H]
\includegraphics[width=0.9\columnwidth]{Figures/StressUpdate}
\caption{{General flowchart} %MDPI: Caption has been extended
 of a FE code, focus on the stresses~computation using an iterative solving~procedure.}
\label{fig:StressUpdate}
\end{figure}
Within the framework of a FE formulation in large deformations such as the one used in thermomechanical modeling of processes, this computation of the stress tensor $\Sig_1$ is to be carried out on all the integration points of each element of the studied structure.
Since the numerical model can include thousands of elements, themselves comprising between $1$ and $8$ integration points depending on the elements used, this stress computation must be as fast as possible in order not to increase the CPU time too much, but~precise enough so that the results are under the physics of the process.
This is even more important if we want to integrate this flow law in an explicit FE code, such as Abaqus/Explicit, for which one second of physical simulation corresponds to several million iterations of these stress computations.
Thus, the~complexity of the ANN, \ie the number of computational steps that must be performed to compute the flow stress as a function of the input variables, is a major parameter in the choice of the neural network.
As an example, for~the model presented in Section~\ref{sec:ANN-traning}, $180$ internal variables, $15$ neurons on the first hidden layer and $7$ neurons on the second hidden layer were listed.
Given the equations described in Sections~\ref{sec:ANN-equations} and~\ref{sec:ANN-derivative}, it will be necessary to compute $22$ exponentials, to~make matrix--vector products of size $15\times3$ and $15\times7$ plus many other numerical operations to compute the flow stress $\sigma^y$ and the $3$ derivatives of it with respect to $\varepsilon^p$, $\mdot\varepsilon$ and $T$.

Implementing the ANN flow law identified above is realized here in a VUHARD subroutine, similarly as proposed by van Rensburg et~al.~\cite{JansenVanRensburg-2012}, used by the Abaqus Explicit FE code in order to allow a user to program the computation of the flow stress $\sigma^y$ and its 3~derivatives as a function of the model input data.
This subroutine is used when calculating the stress tensor $\Sig_1$ at the end of an increment from the stress tensor at the beginning of the increment $\Sig_0$, the~deformations, the~material parameters and the history of the deformation at each finite element integration points, according to the stress integration algorithm based on the radial return method as described in Simo et~al.~\cite{Simo-1998} for the general aspects, Ming et~al.~\cite{Ming-2018} for Abaqus Explicit FE code, or Pantalé et~al.~\cite{Pantale-2004} for the DynELA FE code.
Thus, without~going into too much detail about the stress integration scheme used in finite element codes (the curious reader can refer to~\cite{Ponthot-2002, Ming-2018, Pantale-2004, Liang-2022} for details about this method), Figure~\ref{fig:RadialReturn} shows the location of the VUHARD subroutine used to compute the flow stress $\sigma^y$ and its derivative $\partial\sigma^y/\partial\Gamma$ used in the writing of the two quantities $\gamma(\Gamma)$ and $\gamma^{'}(\Gamma)$ used in the Newton--Raphson solving procedure from the following relation:
\begin{equation}
\frac{d\sigma^{y}}{d\Gamma} = \sqrt{\frac{2}{3}}\left(\frac{\partial\sigma^{y}}{\partial\varepsilon^{p}}+\frac{1}{\Delta t}\frac{\partial\sigma^{y}}{\partial\mdot{\varepsilon}^p}+\frac{\eta\sigma^{y}}{\rho C_{p}}\frac{\partial\sigma^{y}}{\partial T}\right)\label{eq:derivatives},
\end{equation}
where $\Gamma$ is the consistency parameter used in the radial return algorithm as defined by Simo et~al.~\cite{Simo-1998}, $\Delta t$ is the time increment, $\eta$ is the Taylor--Quinney coefficient defining the amount of plastic work converted into heat energy, $C_{p}$ is the specific heat coefficient, $\rho$ is the density of the material and $\partial \sigma^y/\partial\varepsilon^p$, $\partial \sigma^y/\partial\mdot\varepsilon$ and $\partial \sigma^y/\partial T$ are the three derivatives defined in Equation~(\ref{eq:DANN3}).
\begin{figure}[H]
\includegraphics[width=0.6\columnwidth]{Figures/VumatGeneralNewtonRaphson}
\caption{General flowchart of the radial return algorithm to compute the final stress tensor $\Sig_1$.}
\label{fig:RadialReturn}
\end{figure}
\unskip

@@

The mesh of the sample is made with $850$ axis-symmetric quadrilateral finite elements with $4$ nodes and reduced integration (named CAX4R in the Abaqus software) with $50$~elements in the vertical direction and $17$ elements in the radial direction, respectively.
The cylinder is between two rigid surfaces, and the Coulomb friction law with a friction coefficient at the contact surfaces was set to $\mu=0.15$.
The simulation time being fixed at $10~\text{s}$ in~order to reduce the simulation time, considering that an explicit integration scheme is used, a~global mass scaling is used for all simulations.
The VUHARD subroutine is compiled using the GNU gfortran 11.3.0 and linked to the main Abaqus Explicit executable.
All benchmarks tests were solved using Abaqus Explicit 2022 on a Dell XPS 13 laptop running Ubuntu 20.04 64~bits with 16~GiB of RAM and one 4 core i7-10510U Intel Processor.
All computations were performed using the double precision option of Abaqus, with~parallel threads execution on two~cores.

Figure~\ref{fig:peeqContourplot} shows the plastic strain field $\varepsilon^p$ contourplot within the structure at the end of the simulation for both the 3-7-4-1 (left side) and the 3-15-7-1 (right side) flow laws, while Figure~\ref{fig:dpeeqContourplot} shows the temperature field $T$ contourplot for the same~models.
\begin{figure}[H]
\includegraphics[width=0.90\columnwidth]{Figures/peeq}
\caption{{Equivalent plastic} %MDPI: Changes have been done.
 strain $\varepsilon^p$ contourplot for the compression of a cylinder using the 3-7-4-1 (\textbf{left side}) and the 3-15-7-1 (\textbf{right side}) ANN flow~laws.}
\label{fig:peeqContourplot}
\end{figure}
\unskip
\begin{figure}[H]
\includegraphics[width=0.90\columnwidth]{Figures/temp}
\caption{{Temperature} %MDPI: Changes have been done.
 $T$ contourplot for the compression of a cylinder using the 3-7-4-1 (\textbf{left side}) and the 3-15-7-1 (\textbf{right side}) ANN flow~laws.}
\label{fig:dpeeqContourplot}
\end{figure}
Both sides of the figures look more or less the same with some visible differences from the left to the right concerning the shape of the isovalues zones and the maximum value, but~in fact, the two models with the lowest and the highest number of neurons give coherent results concerning for the plastic strain $\varepsilon^p$ and temperature $T$ contourplots.
The maximum plastic strains are concentrated in the center of the specimen with a maximum value of $\varepsilon^p=0.89$ for the 3-7-4-1 model and $\varepsilon^p=0.95$ for the 3-15-7-1 model, which is slightly beyond the limit set by the training data, which varies from $0$ to $0.7$.
As shown by Pantalé et~al.~\cite{Pantale-2021}, the~flow laws defined by neural networks are able to correctly extrapolate the flow stresses $\sigma^y$ when the plastic deformations are higher than at least $150\%$ of the maximum plastic strain used during training.
Concerning the temperature $T$, the~maximum value is around $T=795~\celsius$ and $T=799~\celsius$, which is very close and in accordance to the experiments used for the learning~phase.
